|

### 2021-08-07 prova_knit_finale.Rmd
typo
Funzioni aggiornate (x/2)

The aim of this subsection is to provide a practical illustration of the algorithms discussed in the main chapters. For the sake of simplicity we use for implementations a random walk plus noise model, i.e. the most basic form of linear Gaussian state-space model.  
\begin{align}
y_{t}|x_{t} & \sim N(x_{t},\sigma^{2}) \label{eq:1} \\
x_{t}|x_{t-1} & \sim N(x_{t-1},\tau^{2}) \label{eq:2} \\
x_{0} & \sim N(m_{0},C_{0})
\end{align}
As already mentioned before, in this case the filtering distribution can be computed in closed form solutions using the Kalman filter. However, this toy example will be used also to illustrate more involved filtering strategies described in this work. We believe indeed that it represents a useful starting point to understand the logic of the algorithms which may be eventually replicated when dealing with more complex models. \
The filtering strategy is applied to 50 simulated data. Figure \ref{fig:myfig} shows the simulated true states sequence assuming as data generating process the Equation \eqref{eq:2} with $\tau^{2}=1$ and simulated observed sequence process form Equation \eqref{eq:1} with $\sigma^{2}=1$.
```{r myfig, echo=FALSE, fig.align='center', fig.cap=" Simulated random walk plus noise model", fig.height=4, fig.pos='H', fig.width=8, message=FALSE, warning=FALSE, results=F}

@@

We decide to organise this Appendix as an R package. All the R functions described in this Appendix take as reference a basic specification of Stochastic Volatility Model. 
\begin{align*}
    y_t|x_t & \sim N(0,e^{x_t})\\
    x_t|x_{t-1},\alpha,\beta,\tau^2 & \sim N(\alpha+\beta x_{t-1},\tau^2)
\end{align*}
All the plots and tables of this work concerning stochastic volatility models use this functions.

@@

\code{data} \ \ the observed process. It has to be a vector or a univariate time series.
\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{alpha} \ \ intercept $\alpha$ in the state equation
\
\code{beta} \ \ coefficient $\beta$ in the state equation
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule
\
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
SVPFfun<-function(data,N,m0,C0,alpha,beta,tau,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)

@@

\code{data} \ \ the observed process. It has to be a vector or a univariate time series.
\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{alpha} \ \ intercept $\alpha$ in the state equation
\
\code{beta} \ \ coefficient $\beta$ in the state equation
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule

@@

    xprev<-x
    x<-rnorm(N,alpha+beta*x+(tau^2)/4*((data[t]^2)*exp(-(alpha+beta*x))-2),tau)
    w1<-w*dnorm(data[t],0,exp(x/2))*dnorm(x,alpha+beta*xprev,tau)/
      dnorm(x,mean=alpha+beta*xprev+(tau^2)/4*((data[t]^2)*exp(-(alpha+beta*xprev))-2),tau)

@@

\code{data} \ \ the observed process. It has to be a vector or a univariate time series.
\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{alpha} \ \ intercept $\alpha$ in the state equation
\
\code{beta} \ \ coefficient $\beta$ in the state equation
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule

@@

\code{data} \ \ the observed process. It has to be a vector or a univariate time series.
\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{alpha} \ \ intercept $\alpha$ in the state equation
\
\code{beta} \ \ coefficient $\beta$ in the state equation
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule
\
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
SISfun<-function(data,N,m0,C0,alpha,beta,tau,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)

@@

    xs = rbind(xs,x)
    ws = rbind(ws,w)
    ess =rbind(ess,ESS)
  }
  return(list(xs=xs,ws=ws,ess=ess))
}
```
\
\hrule
\hrule
\
\code{BAPFfun(data,N,m0,C0,alpha,beta,tau,r)}
\
\hrule
\textbf{Arguments}

\code{data} \ \ the observed process. It has to be a vector or a univariate time series.
\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{alpha} \ \ intercept $\alpha$ in the state equation
\
\code{beta} \ \ coefficient $\beta$ in the state equation
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule
\
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
BAPFfun<-function(data,N,m0,C0,alpha,beta,tau,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)

>>>>

The aim of this subsection is to provide a practical illustration of the algorithms discussed in the main chapters. For the sake of simplicity we use for implementations a random walk plus noise model, i.e. the most basic form of linear Gaussian state-space model.  
\begin{align}
y_{t}|x_{t} & \sim N(x_{t},\sigma^{2}) \label{eq:1} \\
x_{t}|x_{t-1} & \sim N(x_{t-1},\tau^{2}) \label{eq:2} \\
x_{0} & \sim N(m_{0},C_{0})
\end{align}
As already mentioned before, in this case the filtering distribution can be computed in closed form solutions using the Kalman filter. However, this toy example will be used also to illustrate more involving filtering strategies described in this work. We believe indeed that it represents a useful starting point to understand the logic of the algorithms which may be eventually replicated when dealing with more complex models. \
The filtering strategy is applied to 50 simulated data. Figure \ref{fig:myfig} shows the simulated true states sequence assuming as data generating process the Equation \eqref{eq:2} with $\tau^{2}=1$ and simulated observed sequence process form Equation \eqref{eq:1} with $\sigma^{2}=1$.
```{r myfig, echo=FALSE, fig.align='center', fig.cap=" Simulated random walk plus noise model", fig.height=4, fig.pos='H', fig.width=8, message=FALSE, warning=FALSE, results=F}

@@

We decide to organise this Appendix as an R package. All the R functions described in this Appendix take as reference a basic specification of Stochastic Volatility Model.\footnote{Actually, the function SVLWfun considers a SV model with an additional hierarchical level, as specified in section XX.}
\begin{align*}
    y_t|x_t & \sim N(0,e^{x_t})\\
    x_t|x_{t-1},\alpha,\beta,\tau^2 & \sim N(\alpha+\beta x_{t-1},\tau^2)
\end{align*}
All the plots and tables of this work concerning stochastic volatility models use this functions.

@@

\code{data} \ \ the observed process. It has to be a vector or a univariate time series.
\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{alpha} \ \ intercept $\alpha$ in the state equation
\
\code{beta} \ \ coefficient $\beta$ in the state equation
\
\code{tau} \ \ square root of $\tau^2$, the conditional variance of the state
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule
\
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
SVPFfun<-function(data,N,m0,C0,alpha,beta,tau,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)

@@

\code{data} \ \ the observed process. It has to be a vector or a univariate time series.
\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{alpha} \ \ intercept $\alpha$ in the state equation
\
\code{beta} \ \ coefficient $\beta$ in the state equation
\
\code{tau} \ \ square root of $\tau^2$, the conditional variance of the state
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule

@@

    xprev<-x
    x<-rnorm(N,alpha+beta*x+(tau^2)/4*((data[t]^2)*exp(-(alpha+beta*x))-2),tau)
    w1<-w*dnorm(data[t],0,exp(x/2))*dnorm(x,alpha+beta*xprev,tau)/dnorm(x,mean=alpha+beta*xprev+(tau^2)/4*((data[t]^2)*exp(-(alpha+beta*xprev))-2),tau)

@@

\code{data} \ \ the observed process. It has to be a vector or a univariate time series.
\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{ealpha} \ \ expected value of the intercept $\alpha$ in the state equation
\
\code{valpha} \ \ variance of the intercept $\alpha$ in the state equation
\
\code{ebeta} \ \ expected value of the coefficient $\beta$ in the state equation
\
\code{vbeta} \ \ variance of the coefficient $\beta$ in the state equation
\
\code{nu} \ \ when divided by 2, first hyperparameter of the Inverse Gamma distribution on $\tau^2$, the conditional variance of the state
\
\code{lambda} \ \ when multiplied by $\nu/2$, second hyperparameter of the Inverse Gamma distribution on $\tau^2$, the conditional variance of the state
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule

@@

\code{data} \ \ the observed process. It has to be a vector or a univariate time series.
\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{alpha} \ \ intercept $\alpha$ in the state equation
\
\code{beta} \ \ coefficient $\beta$ in the state equation
\
\code{tau} \ \ square root of $\tau^2$, the conditional variance of the state
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule
\
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
SISfun<-function(data,N,m0,C0,alpha,beta,tau,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)

@@

```
\
\hrule
\hrule
\
\code{BAPFfun(data,N,m0,C0,alpha,beta,tau,r)}
\
\hrule
\textbf{Arguments}

\code{data} \ \ the observed process. It has to be a vector or a univariate time series.
\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{alpha} \ \ intercept $\alpha$ in the state equation
\
\code{beta} \ \ coefficient $\beta$ in the state equation
\
\code{tau} \ \ square root of $\tau^2$, the conditional variance of the state
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule
\
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
BAPFfun<-function(data,N,m0,C0,alpha,beta,tau,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)
